Forecasting the 2024 US Presidential Election

### 2024-11-03 paper/paper.qmd
Fixed PDF not rendering

\newpage  
# Data {#sec-data}  
## Overview

@@

@@

We sourced the "Presidential General Election Polls" dataset from FiveThirtyEight [@FiveThirtyEight2024] and analyzed it using the statistical programming language R [@citeR].

Our main goal was to clean, organize, and analyze U.S. presidential election polling data to understand voter preferences by state. We selected data from credible pollsters, sorted it by state, and resolved missing or incomplete entries. Where necessary, we merged additional datasets, ensuring that we carefully handled any gaps in the data. The result was a refined dataset ready for analysis. The table below provides a summary of polling data by state, showing voter preferences by party and various demographics. It highlights key trends and differences across states, focusing on party leanings and demographic splits, as illustrated in @fig-predictors-pct-summary.

@@

The correlation matrix in @fig-predictors-party-pct shows how different predictors relate to party preferences. We can see which factors, like education or gender, are more strongly tied to Democratic or Republican leanings. Blue shading indicates stronger Democratic support, while red shows stronger Republican support. Key observations include higher education being linked to Democratic support and lower education favoring Trump in 2020. Minority groups generally lean Democratic, while Caucasian/White voters tend to support Trump. Past voter behavior shows Biden supporters remain Democratic, and Trump supporters disapprove of Biden. Additionally, women lean slightly more Democratic compared to men, who often lean Republican.

@@

# Pollster Methodology Overview (Appendix A) {#sec-appenA}  
Append a detailed review of a chosen pollster's methodology, survey techniques, strengths, and weaknesses.

# Model {#sec-model}
## Model Development

@@

```{r}
#| label: tbl-modelresults
#| tbl-cap: "Forecasting model results based on poll data"
#| echo: false
#| warning: false
#| message: false

# Modify the prediction summary
prediction_summary <- prediction_summary %>%
  select(State, DEM_Mean, REP_Mean) %>%  # Keep only the mean columns
  mutate(Winner = ifelse(DEM_Mean > REP_Mean, "Democrat", "Republican"))  
# Create a formatted table
kable(
  prediction_summary,
  caption = "Prediction Summary for Swing States with Declared Winner",
  col.names = c("State", "Democrat Points", "Republican Points", "Winner"),
  format = "latex", 
  digits = 2
) %>%
  kable_styling(latex_options = c("striped", "hold_position", "condensed"), 
                full_width = FALSE, 
                font_size = 10)
```
*Person B*: Visualize the model's results and include any performance metrics (e.g., RMSE, test/train split).
```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: ""
#| warning: false

@@

## Appendix A: Pollster Methodology Overview
*Person A*: Provide a detailed methodology review for the chosen pollster, including survey design, population sampling, non-response handling, etc.

## Appendix B: Idealized Survey Design for $100K Budget
*Person C*: Develop and describe an ideal survey design for forecasting the election with a $100K budget, and include survey link.

Sampling approach: The sampling approach we will use is stratified random sampling where we divide the population into strata based on.... (age, gender, education level, etc.)
Recruit Respondents: We will recruit respondents using an online survey on Google Forms so that we can minimize the cost and maximize the range of respondents we can reach. We will spend a portion of our budget (specify here) to advertise these surveys and also send out emails, with an additional monetary incentive (specify here) to encourage more participation. 
Data validation: IP address tracking to prevent duplicate responses.
Poll Aggregation: Incorporate Bayesian Inference.

Survey Link: (Short questionnaire asking for demographic questions to be added here)
Copy of Survey:

>>>>

# Data {#sec-data} 
## Overview 
We sourced the "Presidential General Election Polls" dataset from FiveThirtyEight [@FiveThirtyEight2024] and performed an in-depth analysis using the statistical programming language R [@citeR].

@@

Our primary objective was to clean, organize, and analyze U.S. presidential election polling data to provide insights into voter preferences by state. This involved selecting data from reputable pollsters, organizing it by state, and addressing any missing or incomplete entries. Additional datasets were merged where necessary, and after thoroughly addressing missing values, we finalized a polished dataset ready for analysis. Below is a table summarizing the polling data by state, showcasing voter preferences by party and demographic breakdowns. The table highlights key trends and differences across states in terms of party lean and demographic distribution. @fig-predictors-pct-summary-1 and @fig-predictors-pct-summary-2 provides a snapshot of these findings.

@@

@@

```
The correlation matrix in @fig-predictors-party-pct shows how different predictors relate to party preferences. We can see which factors, like education or gender, are more strongly tied to Democratic or Republican leanings. Blue shading indicates stronger Democratic support, while red shows stronger Republican support. Key observations include higher education being linked to Democratic support and lower education favoring Trump in 2020. Minority groups generally lean Democratic, while Caucasian/White voters tend to support Trump. Past voter behavior shows Biden supporters remain Democratic, and Trump supporters disapprove of Biden. Additionally, women lean slightly more Democratic compared to men, who often lean Republican.

@@

# Model {#sec-model}
## Model Development
The goal of our model is to forecast the popular vote outcome of the 2024 US presidential election, focusing specifically on eight swing states. We will choose the predictor variables based on the correlation matrix we created previously (@fig-predictors-party-pct). It is important to note that even if the percentage of people who vote for Trump (representing the Republican Party) or Harris (representing the Democratic Party) is slightly below 50%, this does not necessarily mean that the corresponding candidate will not be elected as the next President of the United States, given the percentage of people who choose not to vote.

```{r}
#| label: fig-hist-for-republican
#| fig-cap: Distribution of Republican Party Supporting Rate
#| echo: false

# Calculate the variable that represents the Republican Party Supporting Rate
# grouped based on various states
data_rep1 <- poll_data %>%
  group_by(state) %>%
  summarise(
    rep_total_pct = round(sum(.data[[column_refs$republican]], na.rm = TRUE), 1)
  )

# Create a histogram based on targeted data variable
ggplot(data = data_rep1, aes(x=rep_total_pct)) +
  geom_histogram(binwidth = 2) +
  labs(title = "Republic Party Supporting Rate Distirbution", x = "Percentage", y = "Count")

The following @fig-hist-for-democratic is the histogram of the Democratic Party Supporting Rate, which describes the distribution of the response variable for the second model. Our final choice is to construct two generalized linear models: one for the percentage that Trump will become the next President and one for Harris. We can then compare the predicted percentages of people voting for each candidate to make a final prediction of who is more likely to be the next President of the United States.

```{r}
#| label: fig-hist-for-democratic
#| fig-cap: Distribution of Democratic Party Supporting Rate
#| echo: false

# Calculate the variable that represents the Republican Party Supporting Rate
# grouped based on various states
data_rep1 <- poll_data %>%
  group_by(state) %>%
  summarise(
    dem_total_pct = round(sum(.data[[column_refs$democrat]], na.rm = TRUE), 1)
  )

# Create a histogram based on targeted data variable
ggplot(data = data_rep1, aes(x=dem_total_pct)) +
  geom_histogram(binwidth = 2) +
  labs(title = "Democratic Party Supporting Rate Distirbution", x = "Percentage", y = "Count")

To decide between using a generalized linear model or a linear model for each of the two models, we checked the normality of the response variables: the percentage of people supporting the Republican Party and the percentage supporting the Democratic Party. According to @fig-hist-for-democratic and @fig-hist-for-republican, the distributions of both response variables do not appear normal. The distribution of the Democratic Party Supporting Rate tends to resemble a Poisson distribution, with most percentages having a similar frequency and an outlier around 67-68 percent. The Republican Party Supporting Rate distribution has a maximum frequency around 50 percent, with similar frequencies around 25-26 percent and 48-49 percent, suggesting a potential violation of the normal distribution assumption.

Therefore, since the response variables are not normally distributed, we will use a generalized linear model for both models to better capture the characteristics of the data.

@@

```{r}
#| label: tbl-modelresults
#| tbl-cap: "Prediction Summary for Swing States with Declared Winner"
#| echo: false
#| warning: false
#| message: false

# Modify the prediction summary
prediction_summary <- prediction_summary %>%
  select(State, DEM_Mean, REP_Mean) %>%  # Keep only the mean columns
  mutate(Winner = ifelse(DEM_Mean > REP_Mean, "Democrat", "Republican"))

# Create a formatted table
kable(
  prediction_summary,
  col.names = c("State", "Democrat Points", "Republican Points", "Winner"),
  format = "latex", 
  digits = 2,
  escape = FALSE
) %>%
  kable_styling(
    latex_options = c("striped", "hold_position", "condensed"), 
    full_width = FALSE, 
    font_size = 10
)

@@

## Appendix A: Pollster Methodology Overview
Emerson College Polling was initially a college polling exercise, and is now an "innovative, nationally-ranked" non-partisan polling center (@AboutUs). For the 2024 USA Presidential Election pollster, the Emerson College Polling defined the population to be "likely voters" of 2024 United States Presidential Election. Emerson College Polling does indicate that the targeted population of "likely voters" is based on "2024 Likely Voter Modeling." However, the methodology for this modelling and the specific model is not indicated, so the specific validity of the targeted population cannot be determine (@Polling).

Emerson College Polling recruits people completing this survey for this national pollster by contacting mobile phones using "MMS-to-Online", "Online Opt-in Panel", and "IVR(Interactive Voice Response)" (@Polling). In the MMS-to-Online approach, the target population were sent text messages with graphics that invite them to take a "screening questionnaire", and those who pass the questionnaire can move on to take the survey. The selected respondents were based on "state voter files provided by Aristotle." The Online Opt-in Panel approach involves respondents from the targeted population being invited to finish the survey by means of an online opt-in panel "provided by CINT" (@AboutUs). Specifically, in the Online Panel approach, selected voters were pre-matched to "L2 voter file data provided by Rep Data" (@Polling). Finally, the IVR approach involves making automatic telephone calls to selected "likely voters." Participants then answer the survey using their "touch-tone telephone". The email approach indicated in Emerson College Polling's official methodology (@AboutUs) is not mentioned in the National Polling, so it can be assumed that this approach is not used in our chosen pollster data.

From the methodology of recruiting people indicated in the previous paragraph, we can see that while the targeted population of the national poll are likely voters, the sampling frame are likely voters that use phones, either mobile or landline. Emerson College Polling chooses 1000 samples from their targeted population, in this case is all likely voters of the 2024 United States Presidential Election, and it is then "weighted by gender, education, race, age, party affiliation, and region based on 2024 likely voter modeling" (@Polling) for each national poll. As most people in the United States use phones by some means, the target population and sampling frame are highly similar, increasing the overall validity of the survey.

In the "MMS-to-Online" and "IVR(Interactive Voice Response)" approaches, random sampling is used, and in the "Online Opt-in Panel" approach, there is no indication of a specific sampling approach (@AboutUs). We can conclude that the general approach used in this specific pollster is random sampling. Random sampling tends to reduce bias, simplify analysis, and is also easy to implement, but at the same time also requires much time and money. In consideration with the sample size of 1000 people compared to the targeted population of likely voters, which contains most of the population of the United States of America, there is a high possibility of existing selection bias, occurring when the participants are not representative of the population.

There is also no specific indication of how non-response is treated, but the results section of national polls, such as the one from September 29 to October 1, 2024, implicates that non-response are eliminated from final recorded results of the survey. Therefore, we cannot ignore the possibility of non-response bias, occurring when non-respondents differ significantly from respondents. For example, such non-response bias occurs when non-respondents are from California, while respondents are from other states except California.

The questionnaire set by Emerson College Polling for this survey, based on the national polls set from September 29 to October 1, 2024, compose mostly of questions with most common choices listed as choices, and also allowing participants to type in their own answer if it is not included in the list of most common choices. The questionnaire is generally written in an objective voice, and all questions only allow participants to make only one choice. All questions in the questionnaire are relatively-common, including ones about ethnicity, age range, region, and education which collects demographical data, and ones that are related to actual presidential-election predictions, including opinions about Joe Biden's performance, the two major parties (Democratic and Republican), whether s/he would vote, and voting inclinations. The only issue observed in the questionnaire is that it is too lengthy, containing more than 20 questions, which may decrease participants' motivations to complete the survey, even after starting it (@Polling).

Therefore, after analyzing the advantages and disadvantages of this pollster about the 2024 United States Presidential Elections by Emerson College Polling, we can conclude that Emerson College Polling can improve their pollster by implementing the following approaches. Emerson can consider increasing the sample size in order to make their survey more representative of the targeted population, explicitly describing their definition of the targeted population "likely voters" and the model used to determine characteristics of the targeted population. Emerson can also consider changing from simple random to stratified sampling, based on different states or other demographical features based on their targeted population model, making their sample more representative of the population, decrease the length of the questionnaire, and also explicitly explain their treatment of non-response, which will increase the validity of the study. Overall, this pollster is still highly valid, and generally reliable for predicting the 2024 United States Presidential Election.

## Appendix B: Idealized Survey Design for $100K Budget

## Introduction

This appendix introduces an idealized methodology and survey to forecast the US presidential election with a budget of $100k. The goal is to use a sample survey that effectively captures public opinion and demographic characteristics.

## Sampling Approach

In order to most accurately represent the sample population of US voters, we will be using a stratified random sampling approach. Stratified random sampling efficiently and fairly samples data from strata that are representative of a population compared to simple random sampling since it also reflects the population of interest’s underrepresented subgroups, ultimately reducing potential bias.

The sample population will be taken from 10,000 eligible US voters represented in strata divided into different demographic factors such as age, gender, race, education level, income level, and political affiliation. This sample size can provide a statistically significant confidence interval for the strata.

## Respondent Recruitment

We will be using 70% of our budget for respondent recruitment, and the remaining 30% for implementing the survey and validating results.

A proportion of our budget of $20,000 will be used for advertising on various social media platforms such as Instagram, Facebook, and Twitter to reach around 10,000 respondents. This online advertising is a good way to reach target respondents based on different demographics we are interested in, especially for younger or hard-to-reach populations.

Since digital ads may result in higher dropout rates and lower response quality, the majority of our budget will be allocated to make up for this limitation. A budget of $40,000 can be used to hire professional survey panels that work with companies that can provide large databases of target respondents that have been pre-screened to match our demographic of interest. 
Lastly, since pre-screened professional survey panels can introduce bias from being frequent survey responders, we will also use incentivized online survey platforms to recruit various participants that fit our demographics of interest. Participants will be offered small monetary incentives with our budget of $10,000, which can greatly increase completion rates and result in better response quality.

## Data Validation

First, we need pre-screening questions to confirm the eligibility of respondents such as eligibility to vote to prevent bots and fake responses. For instance, asking whether they are a US citizen, and then asking if they are eligible to vote in the upcoming US presidential election. If their answers contradict these two questions, we know that their response lacks validity.

Additionally, we have attention checks to identify if respondents are paying attention to and understanding the survey questions. For example, we can ask them if they are paying attention to the question, and if they answer no, we know that they might be rushing or answering the survey carelessly.

Furthermore, we want consistency checks that include similar questions which are paraphrased in different parts of the survey to ensure that respondents are answering consistently.

Lastly, IP address monitoring can be used to prevent duplicate responses from the same person.

## Poll Aggregation

We can collect poll data from various reputable polling sources that have good pollster reliability and are from more recent polls. This way, we can weigh the polls according to their quality. For instance, we would assign higher weights to pollsters that have a high rate of pollster reliability based on past results. We would also give higher weights to more recent polls that better showcase the current opinions of the upcoming US election. Lastly, we would emphasize the weight of larger sample sizes since these provide more accurate representation of the population of interest. We can also incorporate the Bayesian Inference method when updating new polls to adjust prior distributions with new information.

Survey Link: https://forms.gle/vm3gKbsMAZvzakFn8

## Copy of Survey
